---
title: "EDA"
author: "Paul Villanueva"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(igraph)
library(reshape2)
library(scales)
library(tidyverse)
library(GENIE3)

theme_set(theme_light())
```


# Metabolomics data

```{r}
metabolomics <- read.delim("./data/raw/Metabolomics2.txt")
```

# Proteomics

```{r}
proteomics <- read.delim("./data/raw/Proteomics2.txt")
```

To do: calculating correlations.

#  Transcriptomics

```{r}
transcriptomics <- read.delim("./data/raw/Transcriptomics2.txt")
```

### Building GRN

I ran GENIE3 several times for different combinations of samples:

1. PHOTOTROPH1 - PHOTOTROPH3
2. H1 - H3
3. L1 - L3
4. H1 - H3, L1 - L3
5. PHOTOTROPH1 - PHOTOTROPH3, H1 - H3, L1 - L3

I saved the resulting matrices in (unzipped) data/grn_mats/transcriptomics. The name corresponds to the columns I considered. You can read them in with:

```{r}
H1_H2_H3.distmat <- readRDS("./data/grn_mats/transcriptomics/t_H1-H2-H3.rds")
```

And from there, you can make a network by doing:

```{r}
H1_H2_H3.network <- graph_from_adjacency_matrix(H1_H2_H3.distmat)
```

```{r}
plot(H1_H2_H3.network)
```








