---
title: "EDA"
author: "Paul Villanueva"
date: "5/1/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(data.table)
library(doParallel)
library(dplyr)
library(ggplot2)
library(grid)
library(igraph)
library(NetRep)
library(reshape2)
library(scales)
library(tidyverse)
library(GENIE3)
```

`bcb570_utils.R` contains helper functions that we'll use here.

```{r}
source('./src/bcb570_utils.R')

theme_set(theme_light())
```


# Proteomics

Reads in the proteomics data as a dataframe.

```{r}
proteomics <- read.delim("./data/raw/Proteomics2.txt")
```


## H1-H3

Pulls proteomics and creates a seperate dataframe using just the UNIQID and the H and L expression data.

```{r}
H1_H3.treatment_cols <- 8:10
H1_H3.cor <- get_treatment_cor(proteomics, H1_H3.treatment_cols)
```

```{r}
plot_cor_heatmap(H1_H3.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: Histidine mutants")
```


Bootstrapping protein interactions to get qualtiles and thresholds. 

```{r}
H1_H3.boots <- bootstrap_cor(proteomics, treats = H1_H3.treatment_cols, B = 100)
H1_H3.quantiles <- bootstrap_quantiles(H1_H3.boots, 0.10)
```

Visualizing the distribution of the bootstrapped protein-protein correlation. 

```{r}
plot_bootstrap(H1_H3.boots, H1_H3.quantiles)
```

Updated heatmap showing only the significant correlations

```{r}
H1_H3.cor <- threshold_cor_matrix(H1_H3.cor, H1_H3.quantiles)

plot_cor_heatmap(H1_H3.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: Histidine mutants")
```

Set the diagonal to 0 so that they don't get graphed in the network (`diag = FALSE` argument isn't working in `graph_from_adjaceny_matrix` for some reason).

```{r}
write.cor_to_edgelist(H1_H3.cor, "./data/grn_mats/prot_network.H.csv")
```

Creates a protein-protein interaction graph from the (modified) correlation matrix, removes nodes of degree 0, changes the size of each node to be proportional to its degree, and creates a separate attribute for the betweenness of the node.




## Leucine mutant

```{r}
L1_L3.treatment_cols <- 11:13
L1_L3.cor <- get_treatment_cor(proteomics, L1_L3.treatment_cols)
```

```{r}
plot_cor_heatmap(L1_L3.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions in the leucine mutant") + 
  theme(legend.position = "none")
```


Bootstrapping protein interactions to get qualtiles and thresholds. 

```{r}
L1_L3.boots <- bootstrap_cor(proteomics, treats = L1_L3.treatment_cols, B = 100)
L1_L3.quantiles <- bootstrap_quantiles(L1_L3.boots, 0.10)
```

Visualizing the distribution of the bootstrapped protein-protein correlation. 

```{r}
plot_bootstrap(L1_L3.boots, L1_L3.quantiles) + labs(title = "Simulated correlation histogram: Leucine mutant (B = 1000)", x = "Spearman correlation")
```
 

Updated heatmap showing only the significant correlations

```{r}
L1_L3.cor <- threshold_cor_matrix(L1_L3.cor, L1_L3.quantiles)
plot_cor_heatmap(L1_L3.cor, triangle = TRUE) + 
  labs(title = "Significant protein-protein interactions in the leucine mutant")
```
This writes out the protein-protein interaction network for use with Cytoscape.

```{r}
write.cor_to_edgelist(L1_L3.cor, "./data/grn_mats/prot_network.L.csv")
```

## Wildtype


```{r}
WT.treatment_cols <- 14:16
WT.cor <- get_treatment_cor(proteomics, WT.treatment_cols)
```

```{r}
plot_cor_heatmap(WT.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: Wild-type")
```


Bootstrapping protein interactions to get qualtiles and thresholds. 

```{r}
WT.boots <- bootstrap_cor(proteomics, treats = WT.treatment_cols, B = 100)
WT.quantiles <- bootstrap_quantiles(WT.boots, 0.10)
```

Visualizing the distribution of the bootstrapped protein-protein correlation. 

```{r}
plot_bootstrap(WT.boots, WT.quantiles)
```
 

Updated heatmap showing only the significant correlations

```{r}
WT.cor <- threshold_cor_matrix(WT.cor, WT.quantiles)

plot_cor_heatmap(WT.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: Wild-type")
```

```{r}
write.cor_to_edgelist(WT.cor, "./data/grn_mats/prot_network.WT.csv")
```

Creates a protein-protein interaction graph from the (modified) correlation matrix, removes nodes of degree 0, changes the size of each node to be proportional to its degree, and creates a separate attribute for the betweenness of the node.


# All together


```{r}
all.treatment_cols <- 8:16
all.cor <- get_treatment_cor(proteomics, all.treatment_cols)
```

```{r}
plot_cor_heatmap(all.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: All genotypes")
```


Bootstrapping protein interactions to get qualtiles and thresholds. 

```{r}
all.boots <- bootstrap_cor(proteomics, treats = all.treatment_cols, B = 100, method = "pearson")
all.quantiles <- bootstrap_quantiles(all.boots, 0.10)
```

Visualizing the distribution of the bootstrapped protein-protein correlation. 

```{r}
plot_bootstrap(all.boots, all.quantiles) + 
  labs(title = "Simulated correlation histogram: All genotypes",
                                                    subtitle = "n = 1000")
```
 

Updated heatmap showing only the significant correlations

```{r}
all.cor <- threshold_cor_matrix(all.cor, all.quantiles)
plot_cor_heatmap(all.cor, triangle = TRUE) + 
  labs(title = "Protein-protein interactions: Wild-type")
```

This writes out the protein-protein interaction network for use with Cytoscape.

```{r}
write.cor_to_edgelist(all.cor, "./data/grn_mats/prot_network.all.csv")
```


# Networks

```{r}
WT.g <- graph_from_adjacency_matrix(WT.cor, mode = "undirected")
WT.g <- prepare_network(WT.g)
WT.btwn_coms <- cluster_edge_betweenness(WT.g)

H.g <- graph_from_adjacency_matrix(H1_H3.cor, mode = "undirected")
H.g <- prepare_network(H.g)
H.btwn_coms <- cluster_edge_betweenness(H.g)

L.g <- graph_from_adjacency_matrix(L1_L3.cor, mode = "undirected")
L.g <- prepare_network(L.g)
L.btwn_coms <- cluster_edge_betweenness(L.g)
```

Creating color palettes:

```{r}
fine = 1000 
pal = colorRampPalette(c('azure2','darkblue'))

graphCol = pal(fine)[as.numeric(cut(V(L.g)$btwn, breaks = fine))]
```


```{r}
plot(H.btwn_coms,
     H.g,
     vertex.label = NA,
     vertex.size = V(H.g)$btwn * 0.001,
     layout = layout_nicely(H.g),
     edge.arrow.size = 0.1,
     # vertex.color = graphCol,
     # edge.color = ifelse(E(proteomics.g)$weight == 1, 'green', 'red'),
     )
```


```{r}
plot(WT.btwn_coms,
     WT.g,
     vertex.label = NA,
     vertex.size = V(WT.g)$btwn * 0.0015,
     layout = layout_in_circle(WT.g),
     edge.arrow.size = 0.1
     )
```

```{r}
plot(L.btwn_coms,
     L.g,
     vertex.label = NA,
     vertex.size = V(L.g)$btwn * 0.0015,
     layout = layout_nicely(L.g),
     edge.arrow.size = 0.1,
     rescale = T
          # vertex.color = graphCol,
     # edge.color = ifelse(E(proteomics.g)$weight == 1, 'green', 'red'),
     )
```

# Netrep

Data required for NetRep:

* An adjacency matrix whose entries indicate the strength of the relationship between nodes.
* A matrix whose entries contain the correlation coefficient between each pair of nodes in the network.
* A vector containing the module/group label for each node in the network for each discovery dataset.
* Optionally, a “data matrix”, which contains the data used to calculate the correlation structure and infer the network, e.g. gene expression data.

Data things needed:

* `network`: a list of interaction networks, one for each dataset.
* `data`: a list of data matrices used to infer those networks, one for each dataset.
* `correlation`: a list of matrices containing the pairwise correlation coefficients between variables/nodes in each dataset.
* `moduleAssignments`: a list of vectors, one for each discovery dataset, containing the module assignments for each node in that dataset.
* `modules`: a list of vectors, one vector for each discovery dataset, containing the names of the modules from that dataset to run the function on.
* `discovery`: a vector indicating the names or indices to use as the discovery datasets in the network, data, correlation,  moduleAssignments, and modules arguments.
* `test`: a list of vectors, one vector for each discovery dataset, containing the names or indices of the network, data, and correlation argument lists to use as the test dataset(s) for the analysis of each discovery dataset.


## Data prep

```{r}
dim(L1_L3.cor)
dim(WT.cor)
```

```{r}
# L.adj <- as.matrix(get.adjacency(L.g))
# WT.adj <- as.matrix(get.adjacency(WT.g))
networks = list(cohort.1 = WT.cor, 
                cohort.2 = L1_L3.cor)
```

```{r}
ecoli.data <- list(
  cohort.1 = get_expr_data(proteomics, WT.treatment_cols, "UNIQID"),
  cohort.2 = get_expr_data(proteomics, L1_L3.treatment_cols, "UNIQID")
)
```


Module assignments:

```{r}
L.module_labels <- set_names(L.btwn_coms$membership,
                             L.btwn_coms$names)
```

```{r}
ecoli.preservation <- modulePreservation(network = networks,
                                         data = ecoli.data,
                                         correlation = networks,
                                         moduleAssignments = L.module_labels,
                                         nPerm = 5000,
                                         nThreads = 4)
```
```{r}
sigs <- ecoli.preservation$p.values
sigs = cbind(sigs, 
             passes = as.numeric(rowSums(sigs < 0.05, na.rm = TRUE)),
             fails = as.numeric(rowSums(sigs >= 0.05, na.rm = TRUE))
             )
sigs
```

Community 4 has the weakest evidence for preservation. We'll look at that.

```{r}
proteins_of_interest <- communities(L.btwn_coms)[c(4, 15, 17)]
proteins_of_interest
```

```{r}
t <- data.frame(comm = 1:23, sigs) %>% dplyr::select(-fails)
t[is.na(t)] = 1
y.labels = c("Average node contribution", "Density of correlation structure",
             "Concordance of node contribution", 
             "Concordance of weighted degree", 
             "Concordance of correlation structure", "Module coherence", "Average edge weight")
(t.plot <- t %>% dplyr::select(-passes) %>% melt(id.vars = "comm") %>% 
  ggplot(aes(x = comm, y = variable, fill = value)) + 
  geom_tile(color = "white") + coord_equal() + 
  labs(x = "", y = "", fill = "p-value") + 
  scale_y_discrete(labels = y.labels) + 
  scale_fill_gradient(low = "steelblue", high = "gray", space = "Lab") +
  scale_x_continuous(breaks = 1:23) +
  theme(axis.ticks = element_blank(),
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0),
        axis.text.y = element_text(hjust = 1),
        panel.grid = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank()))


```


```{r}
s <- t %>% select(comm, passes) %>% 
  mutate(evidence = ifelse(passes <= 1, 2, 1))
ggplot(s, aes(comm, y = 1)) + 
  geom_point(size = 5, aes(fill = factor(evidence)), color = "black",
             pch = 21) +
  labs(x = "", y = "Evidence") + 
  scale_color_manual(values = c("1" = "White")) + 
  theme(axis.ticks = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_blank(),
        axis.title.y = element_text(angle = 0, vjust = 0.5),
        panel.background = element_blank(),
        panel.border = element_blank(),
        legend.position = "none"
        )
```

