---
title: "Transcriptomics"
author: "Paul Villanueva"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(dplyr)
library(ggplot2)
library(igraph)
library(NetRep)
library(reshape2)
library(scales)
library(tidyverse)
library(GENIE3)
source("./src/bcb570_utils.R")
theme_set(theme_light())
```

# Transcriptomic analysis

Reading in data:

```{r}
transcriptomics <- read.delim("./data/raw/Transcriptomics2.txt")
```

Loading in pre-generated GENIE3 adjacency matrix from transcriptomic data.


```{r}
grn.T_H_P <- readRDS("./data/grn_mats/transcriptomics/t_PROTOTROPH1-PROTOTROPH2-PROTOTROPH3-H1-H2-H3-L1-L2-L3.rds")
```

Making a network from the top 1000 strongest connections.

```{r}
T_H_P.network <- getLinkList(grn.T_H_P, reportMax = 1000) %>% 
  graph.data.frame()
```

Making degree distribution:

```{r}
plot_dd <- function(g){
  g.dd <- data.frame(0:(length(degree_distribution(g)) - 1), 
                     degree_distribution(g, cumulative = FALSE))
  colnames(g.dd) <-c("k", "p_k")
  ggplot(g.dd, aes(x = k, y = p_k)) + geom_bar(stat = "identity") +
    labs(x = "Degree (k)", y = "Proportion of nodes of degree k (p(k))", title = "Degree distribution: k vs. p(k)", 
         subtitle = paste("Graph:", deparse(substitute(g)))) +
    theme(plot.title = element_text(hjust = 0.5), plot.subtitle = element_text(hjust = 0.5))
}
```


```{r}
plot_dd(T_H_P.network) 
```

Seeing how well it fits a power-law:

```{r}
T_H_P.pl_fit <- fit_power_law(degree(T_H_P.network))
T_H_P.pl_fit
```

We see that the exponent of the of exponential curve describing the degree distribution is 2.8, which falls within the typical range for curves that follow a power law. The null hypothesis of the `fit-power-law` function is that the distribution is drawn from a power-law distribution, so the Kolmogorov-Smirnov p-value of 0.67 indicates there is sufficient evidence that the degree distribution of our network follows a powerlaw.

```{r}
power_law_curve <- function(x){
  x^(-T_H_P.pl_fit$alpha)
}
plot_dd(T_H_P.network) +
  stat_function(fun = power_law_curve, color = "red") + 
  ylim(0, 0.6)
```

Writing out the names of the vertices with degree higher than 6 (arbitrary choice) for further analysis

```{r}
V(T_H_P.network)[degree(T_H_P.network) > 6] %>% 
  names() %>% 
  write.table(file = "transcriptomic_network_factors", row.names = FALSE, col.names = FALSE,
              sep = "\n", quote = FALSE)
```

```{r}
T_H_P.network <- prepare_network(T_H_P.network)
```

```{r}
plot(T_H_P.network,
     vertex.label = NA,
     vertex.size = V(T_H_P.network)$btwn * 0.0015,
     layout = layout_nicely(T_H_P.network),
     edge.arrow.size = 0.1,
     rescale = T
          # vertex.color = graphCol,
     # edge.color = ifelse(E(proteomics.g)$weight == 1, 'green', 'red'),
     )
```

```{r}
data.frame(val = V(T_H_P.network)$btwn) %>% 
  ggplot(aes(val)) +
  geom_histogram() + 
  labs(x = "Betweenness score", y = "Counts",
       title = "Histogram of betweenness scores for transcriptomic network")
```




